---
title: "Do Movies That Pass the Bechdel Test Perform Better in the Box Office?"
author: "Zoe Yu"
description: "An analysis of the financial effects of female representation in film"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction

In 1985, American cartoonist Alison Bechdel published a comic strip that would come to define how we construct female representation in film. “I have this rule, see,” explains a character in Bechdel’s story. “I only go to a movie if it satisfies three basic requirements. One, it has to have at least two women in it... who, two, talk to each other about, three, something besides a man.” (Oppenheimer and Ratatouille, unfortunately, don’t pass the Bechdel Test, while Teenage Mutant Ninja Turtles II: The Secret of the Ooze and The Silence of the Lambs do.)

The premise of the Bechdel Test is rooted in the qualitative forces that impact female representation in film — films might not have female characters in the first place, and if there are female characters present, their roles may revolve around their male counterparts or devolve into stereotypes like the “damsel-in-distress” or the “career woman.” In my project, I focus on quantifying this measure in relation to economic performance.

My research question revolves around the relationship between movies that pass or fail the Bechdel Test and how they fare in the box office. This examines both female representation, or the lack thereof, itself, as well as American versus global reactions, audience reception, and success in theaters. I hypothesize that movies that pass the Bechdel Test will be higher-grossing, as passing the Bechdel Test likely lends itself to a more complex, nuanced plotline and story that moviegoers are more inclined to enjoy. I also considered budget as a variable for explaining grossing, as more financial backing and investment into a movie will likely lend itself to a higher quality movie with more resources for promotions, hiring actors, and more.

## The Data

My data is sourced from two datasets. The first dataset originates from a FiveThirtyEight dataset that joins data from BechdelTest.com on over 5,000 movies and their Bechdel pass/fail status and The-Numbers.com, which contains box office financial data for approximately 4,500 movies, to create a new dataset of 1,615 films, their Bechdel pass/fail status, and their financial information. Data from BechdelTest.com is crowdsourced by a group of film enthusiasts who watch movies and score them based on the Bechdel criteria. Data from The-Numbers.com is operated by Nash Information Services and conglomerates movie business data that film studios send in.

The second dataset originates from a second site containing financial information for movies, but with a few distinctions from the variables used by the FiveThirtyEight dataset. Even though the first dataset contained my variables of interest (Bechdel pass/fail status and grossing), FiveThirtyEight did not include movie budgets. By merging the two datasets, I was able to observe both Bechdel pass/fail status and grossing, as well as movie budgets to evaluate its effect. After merging, I tidied the dataframe to select only the variables of interest and to name them based on the measures that I wanted to analyze.

# Merging and Tidying the Dataset

```{r}
#Downloading packages
library(tidyverse)
library(ggplot2)
library(dplyr)

#Reading data files
movies <- read_csv("grossing.csv")
bechdel_test <- read_csv("bechdel.csv")

#Merging and tidying dataframe
bechdel_movies <- movies |>
  inner_join(bechdel_test) |>
  rename(bechdel_test = binary) |>
  select(gross, budget, bechdel_test)
```
My study is cross-sectional, as I'm observing a naturally-assigned treatment and comparing Bechdel pass/fail statuses to movie grossing in one point in time, not before or after the test is failed, like a before-and-after design, or analyzing data over time to draw inferences, like a difference-in-differences design. In looking at this relationship, my independent variable is whether or not the movie passes the Bechdel Test, and my dependent variable is the movie's grossing, as plotted in the histogram below and measured in dollars.

# Visualization of the Dependent Variable

```{r}
#Histogram of dependent variable
bechdel_movies |>
  ggplot(bechdel_movies, mapping = aes(x = gross)) + 
  geom_histogram(color = "black", fill = "darkturquoise") +
  labs(x = "Grossing", y = "Frequency",
       title = "Frequency of Movie Grossing",
       caption = "Gov 50 Final Project") +
  theme_classic()+
  scale_x_log10(labels = scales::dollar_format())
```

## Results

# Visualization of Bechdel Pass/Fail Status and Grossing
```{r}
#Boxplot of grossing vs. Bechdel pass/fail status
bechdel_plot <- bechdel_movies |>
  ggplot(aes(x = bechdel_test, y = gross, fill = bechdel_test, na.rm = TRUE)) +
  geom_boxplot() +
  geom_smooth() +
  labs(x = "Bechdel Pass/Fail Status", y = "Grossing",
       title = "Grossing vs. Bechdel Pass/Fail Status",
       caption = "Gov 50 Final Project") +
  theme_classic() + 
  theme(legend.position = "none")

bechdel_plot
```
Based on the boxplot above, the distribution of moving grossing appears to be relatively uniform across Bechdel pass/fail statuses. Movies that passed the Bechdel Test, however, have a lower minimum, median, 25th, and 75th percentiles compared to movies that failed the Bechdel Test. In addition, movies that passed the Bechdel Test have a more pronounced outlier that far exceeds the median grossing threshold for both statuses.

# Correlation Between Bechdel Pass/Fail Status and Grossing
```{r}
#Coding y-variable as numeric
bechdel_movies <- bechdel_movies |>
  mutate(bechdel_binary = if_else(bechdel_test == "FAIL", 0, 1))

#Calculating correlation between Bechdel pass/fail status and grossing
cor(bechdel_movies$gross, bechdel_movies$bechdel_binary, use = "pairwise")

#Plotting correlation
bechdel_movies |>
  ggplot(aes(x = gross, y = bechdel_binary)) +
  geom_jitter() + 
  geom_smooth(method = "lm", color = "darkturquoise") +
  theme_classic() +
  labs(x = "Grossing", y = "Bechdel Pass/Fail Status",
       title = "Correlation Between Bechdel Pass/Fail Status and Grossing",
       caption = "Gov 50 Final Project")
```
To further explore the relationship between Bechdel pass/fail status and grossing, I calculated the correlation between the two variables, finding a correlation of -0.071. The weak negative correlation suggests that higher-grossing movies are less likely to pass the Bechdel Test. After plotting the line of best fit, the scattered data points and best fit line reaffirm the conclusion that there is a weak negative correlation.

# Regression Models
```{r}
bechdel_fit <- lm(gross ~ bechdel_binary, bechdel_movies)

var_labels <- c(
"(Intercept)" = "Intercept",
"bechdel_binary" = "Bechdel Pass/Fail Status"
)
modelsummary::modelsummary(bechdel_fit,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           coef_map = var_labels,
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))
```
To evaluate the statistical significance of this relationship, I performed a simple regression of Bechdel pass/fail status on movie grossing. Based on an alpha level of 0.05 or even 0.01, with a p-value of <0.001 that's lower than the alpha, passing or failing the Bechdel Test has a highly statistically significant effect on grossing. There is a less than 0.001 percent chance that the observed differences are due to chance alone, rejecting the null hypothesis that there is no effect between Bechdel pass/fail status and grossing.

```{r}
bechdel_budget_fit <- lm(gross ~ bechdel_binary + budget, bechdel_movies)

var_labels <- c(
"(Intercept)" = "Intercept",
"bechdel_binary" = "Bechdel Pass/Fail Status",
"budget" = "Budget"
)
modelsummary::modelsummary(bechdel_budget_fit,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           coef_map = var_labels,
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))
```
There may, however, be other variables like budget at play. To analyze the statistical significance of budget as a control, I calculated a multiple regression. With a p-value of 0.390, passing or failing the Bechdel Test did not have a statistically significant effect. Budget, on the other hand, had a high statistical significance with a p-value of <0.001, which is significantly lower than the 0.05 or 0.01 alpha value threshold. The regression indicates that other variables unrelated to Bechdel pass/fail status may be more significant in determining grossing. I failed to reject the null hypothesis that there is no effect between Bechdel pass/fail status and grossing.

## Conclusion
Although the weak negative correlation indicates that there may be a relationship between passing or failing the Bechdel Test and movie grossing, the regression does not support my hypothesis that movies that pass the Bechdel Test will be higher-grossing. In analyzing Bechdel pass/fail status alone, the simple regression indicated statistical significance; however, when analyzing budget as a control, there was no statistical significance, indicating that other variables like budget may contribute to determining grossing. Limitations of this project include other variables besides female representation or even the examined budget as a control that may influence movie grossing, such as the presence of a popular actor or contexts like the Barbenheimer trend, that were not included in the dataset. With more time and money, I would analyze other measures of female representation, including the presence of a female director or substantially-female cast, in relation to the variables outlined above, to further examine how female representation impacts how movies perform in the box office.

## References
(n.d.). Bechdel Test Movie List. https://bechdeltest.com/

Abdelfatah, R., & Arablouei, R. (2023, April 5). What is the Bechdel test? A shorthand for measuring representation in movies. NPR. https://www.npr.org/2023/04/05/1168116147/what-is-the-bechdel-test-a-shorthand-for-measuring-representation-in-movies

From 'Oppenheimer' to 'Avatar': 13 Modern Movies That Surprisingly Fail the Bechdel Test. (2023, August 21). Collider. https://collider.com/modern-movies-bechdel-test-fail/

Hickey, W. (2014, April 1). The Dollar-And-Cents Case Against Hollywood's Exclusion of Women. FiveThirtyEight. https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/

Soderbergh, S. (2023, June 14). The Bechdel Test: 18 Surprising Movies That Pass. Jezebel. https://jezebel.com/movies-bechdel-test-surprising-1850537372
